---
title: "Some R code"
author: "JM M"
date: "January 20, 2016"
output: html_document
---

Here I present some R coding. 
```{r}
ecoData=read.csv("ECO.csv",stringsAsFactors=F)
head(ecoData)
str(ecoData)
```

This is a **summary** of the *"Index of Economic Freedom"*:

```{r, echo=FALSE}
summary(ecoData$eco)
```


## How is the Economic Freedom distributed in the world?

I will add the data from the Index of Economic freedom as another attribute of the countries in the map of the world. The map was downloaded from [this link](http://thematicmapping.org). These are the steps I followed.

### Merge the data of my CSV with the DBF using a common field:
This is not a common merge. You need to do the merge, keeping all the values of the map, and keeping in the same order. So a **sorter** variable is added to re order the map data before saving. 
```{r,warning=FALSE,message=FALSE,cache=TRUE}
library(foreign)
DBF=file.path("map","TM_WORLD_BORDERS-0.3.dbf")
mapData=read.dbf(DBF)
names(mapData)
ecoData=ecoData[,c(5,2)] # just the iso3 and eco value
ecoData=ecoData[complete.cases(ecoData),] # no NAs
mapData$sorter <- as.numeric(rownames(mapData)) # extra var to re order
mapData=merge(mapData,ecoData,by.x="ISO3", by.y="iso3",all.x = T)
mapData=mapData[order(mapData$sorter),] # reordering
rownames(mapData)=NULL
write.dbf(mapData,DBF) #saving updated file
```

### Time to plot the Index:
Now we can plot, the <b style='color:green'>greener</b> the better, the <b style='color:red'>redder</b> the worse. Black is missing value.

```{r,warning=FALSE,message=FALSE,fig.width=12,fig.height=6}
library(maptools)
SHP=file.path("map","TM_WORLD_BORDERS-0.3.shp")
worldmap=readShapeSpatial(SHP) 
library(ggplot2)
worldmapGG <- fortify(worldmap, region = "ISO3") 
map= ggplot(data = worldmap@data,
            aes(map_id = ISO3)) + 
     geom_map(aes(fill = worldmap@data$eco), 
              map = worldmapGG) + 
     expand_limits(x = worldmap@data$LON, 
                   y = worldmap@data$LAT) + 
     scale_fill_continuous(low='red', high='green', 
                           na.value = "white") + 
     guides(fill=FALSE) #no AREA legend

map
```



